---
title: "Probiotic evolution"
author: "Roy Moger-Reischer and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin = 2.54cm
---

# Biomass of ancestral and evolved strains on different resources (Fig. 1)

## Clear workspace and set directory

```{r setup}
rm(list=ls())
knitr::opts_knit$set(root.dir=normalizePath("~/GitHub/Bifidobacterium"))
getwd()
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("knitr")
require("extrafont")
require("ggrepel");
require("gridExtra")
require("contrast")
require("vegan")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) (sd(x)/mean(x))*100
```


# Load data

```{r}
mutpopdf <- read.csv("./Data/num.muts.per.signif.gene.per.popn.csv")
gxp.all.raw <- read.csv("./Data/Bifidobacterium_labels_gxp.csv") 

gxp.signif.raw <- read.csv("./Data/Bifidobacterium_labels_gxp_signif.only.csv")
gxp.signif.raw$sample <- as.factor(gxp.signif.raw$sample)
gxp.signif.raw$treatment <- as.factor(gxp.signif.raw$treatment)
levels(gxp.signif.raw$treatment)[levels(gxp.signif.raw$treatment)=="F"] <- "DP0"
levels(gxp.signif.raw$treatment)[levels(gxp.signif.raw$treatment)=="P"] <- "DP5"
levels(gxp.signif.raw$treatment)[levels(gxp.signif.raw$treatment)=="X"] <- "DP24"
gxp.signif <- as_tibble(gxp.signif.raw)
gxp.signif <- as.matrix(gxp.signif[,3:ncol(gxp.signif)])

gxp.all.raw$sample <- as.factor(gxp.all.raw$sample)
gxp.all.raw$treatment <- as.factor(gxp.all.raw$treatment)
levels(gxp.all.raw$treatment)[levels(gxp.all.raw$treatment)=="F"] <- "DP0"
levels(gxp.all.raw$treatment)[levels(gxp.all.raw$treatment)=="P"] <- "DP5"
levels(gxp.all.raw$treatment)[levels(gxp.all.raw$treatment)=="X"] <- "DP24"
gxp.all <- as_tibble(gxp.all.raw)
gxp.all <- as.matrix(gxp.all[,3:ncol(gxp.all)])
```

# Statistics 

## 1. Does number of putatively adaptive mutation differ among diets?

```{r}
mutnum.aov <- aov(mutpopdf$num.muts ~ mutpopdf$Evolution_treatment)
summary(mutnum.aov)
# Answer: No

# Mean and SEM of putatively adaptive mutations by diet type
mutnum.summary <- mutpopdf %>%
  group_by(Evolution_treatment) %>%
  summarize(mean = mean(num.muts, na.rm = TRUE), sem = sem(num.muts))

# Diet        mean   sem
# DP0         5.28 0.291
# DP5         6.03 0.808
# DP24        6.20 0.579
```

## 2. Does compositions of putatively adaptive mutation differ among diets?

```{r} 
# First, test using PERMANOVA
mut.comp.adonis <- adonis2(gxp.signif ~ gxp.signif.raw$treatment, 
                   method = "bray", permutations = 9999)

#             Df SumOfSqs    R2      F      Pr(>F)    
# Diet        2   1.6579  0.26526   3.7907  1e-04 ***
# Residual   21   4.5922  0.73474                  
# Total      23   6.2500  1.00000      

# Answer: Yes
```


```{r}
# Second, conduct PCoA and visualize ordination

gxp.signif.dist <- vegdist(gxp.signif, method = 'bray', upper = TRUE, diag = TRUE)
gxp.signif.dist <- as_tibble(data.matrix(gxp.signif.dist))

# Run PCoA and quantify explained variance
pcoa.eig <- cmdscale(gxp.signif.dist, eig = TRUE, k = 3)
explainvar1 <- round(pcoa.eig$eig[1] / sum(pcoa.eig$eig), 3) * 100 # 29.4 %
explainvar2 <- round(pcoa.eig$eig[2] / sum(pcoa.eig$eig), 3) * 100 # 18.2 %
explainvar3 <- round(pcoa.eig$eig[3] / sum(pcoa.eig$eig), 3) * 100 # 17.1 %
sum.eig <- sum(explainvar1, explainvar2, explainvar3) # 64.7 %

# Add sample and treatment IDs
gxp.signif.pcoa <- as.data.frame(pcoa.eig[1])
gxp.signif.pcoa$treatment <- gxp.signif.raw$treatment
row.names(gxp.signif.pcoa) <- gxp.signif.raw$sample
gxp.signif.pcoa$sample <- gxp.signif.raw$sample
names(gxp.signif.pcoa)[1:3] <- c('PCo1', 'PCo2', 'PCo3')

# Create ordination (Fig. S3)

png(filename="./Figures/Fig.S3a.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

plot(gxp.signif.pcoa[ ,1], gxp.signif.pcoa[ ,2],
     ylim = c(-0.55, 0.55), xlim = c(-0.75, 0.4),
     xlab = "", ylab = "", xaxt = "n", yaxt = "n",
     pch = 22, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1,
     axes = FALSE)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-0.5", "0.0", "0.5"),
     at = c(-0.5, 0, 0.5))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(-0.5, 0, 0.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-0.5", "-0.25", "0.0", "0.25"), at = c(-0.5, -0.25, 0, 0.25))
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(-0.5, -0.25, 0, 0.25), labels = F)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add axis labels
mtext(expression(paste("PCoA 1 (29.4 %)", sep = "")), side = 1,
      outer = TRUE, cex = 1.5, line = -3.0)
mtext(expression(paste("PCoA 2 (18.2 %)", sep = "")), side = 2,
      outer = TRUE, cex = 1.5, line = -3.0)

# Subset PCoA scores by strain
signif.F <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP0"), ]
signif.P <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP5"), ]
signif.X <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP24"), ]

# Plot points by strain
points(signif.F[ ,1], signif.F[ ,2], pch = 21,
       cex = 1, col = "tomato1", bg = "tomato1", lwd = 2)

points(signif.P[ ,1], signif.P[ ,2], pch = 21,
       cex = 1, col = "springgreen4", bg = "springgreen4", lwd = 2)   

points(signif.X[ ,1], signif.X[ ,2], pch = 21,
       cex = 1, col = "cornflowerblue", bg = "cornflowerblue", lwd = 2)

# Add hulls
tomato1_rgb <- col2rgb("tomato")/255 
springgreen4_rgb <- col2rgb("springgreen4")/255 
cornflowerblue_rgb <- col2rgb("cornflowerblue")/255 

ordihull(cbind(gxp.signif.pcoa$PCo1, gxp.signif.pcoa$PCo2), 
    groups = gxp.signif.pcoa$treatment, lwd = 2, lty = 0, draw = "polygon", 
    col = c(rgb(tomato1_rgb[1], tomato1_rgb[2], tomato1_rgb[3], 0.5),
    rgb(springgreen4_rgb[1], springgreen4_rgb[2], springgreen4_rgb[3], 0.5),
    rgb(cornflowerblue_rgb[1], cornflowerblue_rgb[2], cornflowerblue_rgb[3], 0.5)), 
    label = FALSE)

text(-0.5, 0.43, labels = "DP0", cex = 1.3, col = "tomato1")
text(0.125, 0.42, labels = "DP5", cex = 1.3, col = "springgreen4")
text(0.125, -0.42, labels = "DP24", cex = 1.3, col = "cornflowerblue")

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./Figures/Fig.S3a.png")
grid.raster(img)

# Create ordination (Fig. 3)

png(filename="./Figures/Fig.3a.png",
    width = 1200, height = 1200, res = 96*2) 

plot.new()
par(mar = c(7, 7, 5, 7))

plot(gxp.signif.pcoa[ ,1], gxp.signif.pcoa[ ,3],
     ylim = c(-0.55, 0.55), xlim = c(-0.75, 0.4),
     xlab = "", ylab = "", xaxt = "n", yaxt = "n",
     pch = 22, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1,
     axes = FALSE)

# Add axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-0.5", "0.0", "0.5"),
     at = c(-0.5, 0, 0.5))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at=c(-0.5, 0, 0.5), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-0.5", "-0.25", "0.0", "0.25"), at = c(-0.5, -0.25, 0, 0.25))
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(-0.5, -0.25, 0, 0.25), labels = F)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add axis labels
mtext(expression(paste("PCoA 1 (29.4 %)", sep = "")), side = 1,
      outer = TRUE, cex = 1.5, line = -3.0)
mtext(expression(paste("PCoA 3 (17.1 %)", sep = "")), side = 2,
      outer = TRUE, cex = 1.5, line = -3.0)

# Subset PCoA scores by strain
signif.F <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP0"), ]
signif.P <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP5"), ]
signif.X <- gxp.signif.pcoa[which(gxp.signif.pcoa$treatment == "DP24"), ]

# Plot points by strain
points(signif.F[ ,1], signif.F[ ,3], pch = 21,
       cex = 1, col = "tomato1", bg = "tomato1", lwd = 2)

points(signif.P[ ,1], signif.P[ ,3], pch = 21,
       cex = 1, col = "springgreen4", bg = "springgreen4", lwd = 2)   

points(signif.X[ ,1], signif.X[ ,3], pch = 21,
       cex = 1, col = "cornflowerblue", bg = "cornflowerblue", lwd = 2)

# Add hulls
ordihull(cbind(gxp.signif.pcoa$PCo1, gxp.signif.pcoa$PCo3), 
    groups = gxp.signif.pcoa$treatment, lwd = 2, lty = 0, draw = "polygon", 
    col = c(rgb(tomato1_rgb[1], tomato1_rgb[2], tomato1_rgb[3], 0.5),
    rgb(springgreen4_rgb[1], springgreen4_rgb[2], springgreen4_rgb[3], 0.5),
    rgb(cornflowerblue_rgb[1], cornflowerblue_rgb[2], cornflowerblue_rgb[3], 0.5)), 
    label = FALSE)

text(-0.44, 0.37, labels = "DP0", cex = 1.3, col = "tomato1")
text(0.0, 0.47, labels = "DP5", cex = 1.3, col = "springgreen4")
text(0.2, -0.35, labels = "DP24", cex = 1.3, col = "cornflowerblue")

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("./Figures/Fig.3a.png")
grid.raster(img)
```

## 3. Does the variance of multivariate composiiton of differ between diets?

```{r}
gd <- vegdist(gxp.signif, method = 'bray', upper = TRUE, diag = TRUE)
compare.variance <- betadisper(d = gd, group = c("F", "F", "F", "F", "F", "F", "F", "F",
          "G-F5","G-F5","G-F5","G-F5","G-F5","G-F5","G-F5", "G-F5", 
          "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24"))
c.v.result <- anova(compare.variance)
c.v.result
TukeyHSD(compare.variance)

# Answer: Yes
# DP0 has more variance than DP5 or DP24
```

# Do indicator gene analysis
```{r}
trts.i <- c("F", "F", "F", "F", "F", "F", "F", "F",
          "G-F5", "G-F5", "G-F5", "G-F5", "G-F5", "G-F5", "G-F5", "G-F5", 
          "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24", "G-F24")
gxp.i <- gxp.signif
indvals <- multipatt(x = gxp.i, cluster=trts.i, func = "IndVal.g", control = how(nperm = 999))
summary(indvals, indvalcomp = TRUE)
# Component A: "specificity" or "positive predictive value" of the species as indicator of the site group. 
# It is the sample estimate of the probability that the surveyed site belongs to the target site group given the fact that the species has been found.
# Component B: "Fidelity" or "sensitivity" of the species as indicator of the target site group. It is is sample estimate of the probability of finding the species in sites belonging to the site group.

gxpcomb <- combinespecies(gxp.i, max.order = 2)$XC
dim(gxpcomb)
indvalcomb <- multipatt(x = gxpcomb, cluster = trts.i, duleg = TRUE,control = how(nperm = 999))
summary(indvalcomb, indvalcomp = TRUE)

####This is using the whole matrix instead of specifically the significant genes:
gxpa.i <- gxp.all
indval <- multipatt(x = gxpa.i, cluster = trts.i,func = "IndVal.g", control = how(nperm = 999))
summary(indval, indvalcomp = TRUE)

gxpacomb <- combinespecies(gxpa.i, max.order = 2)$XC
dim(gxpacomb)
indvalacomb <- multipatt(x = gxpacomb, cluster = trts.i, duleg = TRUE,control = how(nperm = 999))
summary(indvalacomb, indvalcomp = TRUE)
#I just commented out the 2 above lines so that the whole script runs faster when I run it to declare my variables
```

