---
title: "Probiotic evolution"
author: "Roy Moger-Reischer and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin = 2.54cm
---

# Metabolites of ancestral and evolved strains on different resources (Fig. 4)

# Clear workspace and set directory

```{r setup}
rm(list = ls())
knitr::opts_knit$set(root.dir=normalizePath("~/GitHub/Bifidobacterium"))
getwd()
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("knitr")
require("extrafont")
require("ggrepel");
require("gridExtra")
require("contrast")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) (sd(x)/mean(x))*100
```

# Load data

```{r}
hplc <- read.csv("~/GitHub/Bifidobacterium/data/metabolite.flux.data.csv")
hplc$condition_short <- factor(hplc$condition_short, levels = c("GF","F"))
hplc$evo <- factor(hplc$evo, levels = c("F","P","X","Anc"))
hplc$evo_plot <- recode(hplc$evo, "F" = "DP0", "P" = "DP5", "X" = "DP24", 
                        "Anc" = "Ancestor")
hplc$evo_plot <- factor(hplc$evo_plot, levels = c("DP0", "DP5", "DP24", 
                        "Ancestor"))

# Remove the mixed glucose-fructose experimental data
hplc <- hplc %>%
  filter(condition_short !="GF")
```

# Statistics 

## 1. Calculate metabolite fluxes on ancestor 

```{r}
# Filter ancestral population
anc <- hplc %>%
  filter(evo_trt == "Anc")

# Ancestral fructose flux
anc.summary.F <- anc %>%
  summarize(mean = mean(flux.fru.pos, na.rm = TRUE), sem = sem(flux.fru.pos))

# Ancestral acetate flux
anc.summary.A <- anc %>%
  summarize(mean = mean(flux.ace.0, na.rm = TRUE), sem = sem(flux.ace.0))

# Ancestral lactate flux
anc.summary.L <- anc %>%
  summarize(mean = mean(flux.lact.0, na.rm = TRUE), sem = sem(flux.lact.0))
```

## 2. Metabolite flux differ among strains evovled on different resources? 

```{r}
# Filter evolved populations
evo <- hplc %>%
  filter(evo_trt !="Anc")

# Fructose flux: 

evo.summary.F <- evo %>%
  group_by(evo_plot) %>%
  summarize(mean = mean(flux.fru.pos, na.rm = TRUE), sem = sem(flux.fru.pos))
evo.summary.F <- as.data.frame(evo.summary.F)

# Summary: DP0-evolved > DP5-evolved
evo.aov.F <- aov(flux.fru.pos ~ evo_plot, data = evo)
summary(evo.aov.F)
#              Df Sum Sq Mean Sq F value  Pr(>F)   
# evo_plot     2  109.0   54.52   4.201 0.0292 *
# Residuals   21  272.5   12.98    

TukeyHSD(evo.aov.F)
#              diff        lwr        upr     p adj
# DP5-DP0  -5.139350 -9.679038 -0.5996615 0.0247383
# DP24-DP0 -1.773765 -6.313453  2.7659234 0.5942214
# DP24-DP5  3.365585 -1.174104  7.9052733 0.1725517

#Acetate flux: 

evo.summary.A <- evo %>%
  group_by(evo_plot) %>%
  summarize(mean = mean(flux.ace.0, na.rm = TRUE), sem = sem(flux.ace.0))
evo.summary.A <- as.data.frame(evo.summary.A)

# summary: marginal significance overall;
#          no differences with Tukey
evo.aov.A <- aov(flux.ace.0 ~ evo_plot, data = evo)
summary(evo.aov.A)
#             Df Sum Sq Mean Sq F value  Pr(>F)   
# evo_plot     2  149.9   74.94   2.888  0.078 .
# Residuals   21  545.0   25.95       

TukeyHSD(evo.aov.A)
#              diff         lwr       upr     p adj
# DP5-DP0  -5.297965753 -11.718342  1.122410 0.1183326
# DP24-DP0  0.006701859  -6.413674  6.427078 0.9999962
# DP24-DP5  5.304667612  -1.115708 11.725043 0.1177591

#Lactate flux: 

evo.summary.L <- evo %>%
  group_by(evo_plot) %>%
  summarize(mean = mean(flux.lact.0, na.rm = TRUE), sem = sem(flux.lact.0))
evo.summary.L <- as.data.frame(evo.summary.L)

# summary: no differences;

evo.aov.L <- aov(flux.lact.0 ~ evo_plot, data = evo)
summary(evo.aov.L)
#            Df Sum Sq Mean Sq F value Pr(>F)  
# evo_plot     2   61.3   30.63   1.458  0.255
# Residuals   21  441.1   21.00    

TukeyHSD(evo.aov.L)
#               diff      lwr        upr     p adj
# DP5-DP0  -2.9962208 -8.772062 2.779621 0.4067360
# DP24-DP0  0.6818265 -5.094015 6.457668 0.9524774
# DP24-DP5  3.6780473 -2.097794 9.453889 0.2655777
```

## 3. Did evolution lead to altered metabolite fluxes? 

```{r}
# Fructose

# Subtract flux of each evolved replicate from flux of average ancestor

# Anc0 - Evo0
evo.DP0.F <- evo %>%
  filter(evo_plot == "DP0") %>%
  pull(flux.fru.pos)
delta.DP0.F <- unlist(evo.DP0.F - as.numeric(unlist(anc.summary.F[1,2])))

# Anc5 - Evo5
evo.DP5.F <- evo %>%
  filter(evo_plot == "DP5") %>%
  pull(flux.fru.pos)
delta.DP5.F <- unlist(evo.DP5.F - as.numeric(unlist(anc.summary.F[1,2])))

# Anc24 - Evo24
evo.DP24.F <- evo %>%
  filter(evo_plot == "DP24") %>%
  pull(flux.fru.pos)
delta.DP24.F <- unlist(evo.DP24.F - as.numeric(unlist(anc.summary.F[1,2])))

# Create data frame for analysis
delta.F <- data.frame(flux = c(delta.DP0.F, delta.DP5.F, delta.DP24.F),
  comp = c(rep("DP0", length(delta.DP0.F)), rep("DP5", length(delta.DP5.F)), 
          rep("DP24", length(delta.DP24.F))))
delta.F$comp <- factor(delta.F$comp, levels = c("DP0", "DP5", "DP24"))

# Summarize deltas
delta.summary.F <- delta.F %>%
  group_by(comp) %>%
  summarize(mean = mean(flux, na.rm = TRUE), sem = sem(flux))

#   comp   mean   sem
# 1 DP0    9.58 1.56 
# 2 DP5    4.44 1.20 
# 3 DP24   7.80 0.994

# GLM without intercept: test whether groups of delta fluxes are different from zero
delta.fit1 <- lm(flux ~ comp - 1, data = delta.F)
summary(delta.fit1)

# Coefficients:
#       Estimate Std. Error t value Pr(>|t|)   
# compDP0     9.578      1.274   7.521 2.18e-07 ***
# compDP5     4.439      1.274   3.486  0.00221 ** 
# compDP24    7.805      1.274   6.128 4.42e-06 ***

# Residual standard error: 3.602 on 21 degrees of freedom
# Multiple R-squared:  0.835,	Adjusted R-squared:  0.8114 
# F-statistic: 35.42 on 3 and 21 DF,  p-value: 2.119e-08

# Conclusion: fructose flux increased on all resource sizes during evolution


# Acetate

# Subtract flux of each evolved replicate from flux of average ancestor

# Anc0 - Evo0
evo.DP0.A <- evo %>%
  filter(evo_plot == "DP0") %>%
  pull(flux.ace.0)
delta.DP0.A <- unlist(evo.DP0.A - as.numeric(unlist(anc.summary.A[1,2])))

# Anc5 - Evo5
evo.DP5.A <- evo %>%
  filter(evo_plot == "DP5") %>%
  pull(flux.ace.0)
delta.DP5.A <- unlist(evo.DP5.A - as.numeric(unlist(anc.summary.A[1,2])))

# Anc24 - Evo24
evo.DP24.A <- evo %>%
  filter(evo_plot == "DP24") %>%
  pull(flux.ace.0)
delta.DP24.A <- unlist(evo.DP24.A - as.numeric(unlist(anc.summary.A[1,2])))

# Create data frame for analysis
delta.A <- data.frame(flux = c(delta.DP0.A, delta.DP5.A, delta.DP24.A),
  comp = c(rep("DP0", length(delta.DP0.A)), rep("DP5", length(delta.DP5.A)), 
          rep("DP24", length(delta.DP24.A))))
delta.A$comp <- factor(delta.A$comp, levels = c("DP0", "DP5", "DP24"))

# Summarize deltas
delta.summary.A <- delta.A %>%
  group_by(comp) %>%
  summarize(mean = mean(flux, na.rm = TRUE), sem = sem(flux))

#   comp   mean   sem
# 1 DP0    15.3  1.93
# 2 DP5    10.0  1.56
# 3 DP24   15.3  1.89

# GLM without intercept: test whether groups of delta fluxes are different from zero
delta.fit2 <- lm(flux ~ comp - 1, data = delta.A)
summary(delta.fit2)

# Coefficients:
#       Estimate Std. Error t value Pr(>|t|)   
# compDP0    15.305      1.801   8.498 3.09e-08 ***
# compDP5    10.007      1.801   5.556 1.63e-05 ***
# compDP24   15.312      1.801   8.501 3.07e-08 ***

# Residual standard error: 5.094 on 21 degrees of freedom
# Multiple R-squared:  0.8931,	Adjusted R-squared:  0.8778 
# F-statistic: 58.45 on 3 and 21 DF,  p-value: 2.302e-10

# Conclusion: acetate flux increased on all resource sizes during evolution



# Lactate

# Subtract flux of each evolved replicate from flux of average ancestor

# Anc0 - Evo0
evo.DP0.L <- evo %>%
  filter(evo_plot == "DP0") %>%
  pull(flux.lact.0)
delta.DP0.L <- unlist(evo.DP0.L - as.numeric(unlist(anc.summary.L[1,2])))

# Anc5 - Evo5
evo.DP5.L <- evo %>%
  filter(evo_plot == "DP5") %>%
  pull(flux.lact.0)
delta.DP5.L <- unlist(evo.DP5.L - as.numeric(unlist(anc.summary.L[1,2])))

# Anc24 - Evo24
evo.DP24.L <- evo %>%
  filter(evo_plot == "DP24") %>%
  pull(flux.lact.0)
delta.DP24.L <- unlist(evo.DP24.L - as.numeric(unlist(anc.summary.L[1,2])))

# Create data frame for analysis
delta.L <- data.frame(flux = c(delta.DP0.L, delta.DP5.L, delta.DP24.L),
  comp = c(rep("DP0", length(delta.DP0.L)), rep("DP5", length(delta.DP5.L)), 
          rep("DP24", length(delta.DP24.L))))
delta.L$comp <- factor(delta.L$comp, levels = c("DP0", "DP5", "DP24"))

# Summarize deltas
delta.summary.L <- delta.L %>%
  group_by(comp) %>%
  summarize(mean = mean(flux, na.rm = TRUE), sem = sem(flux))

#   comp   mean   sem
# 1 DP0    5.64  2.03
# 2 DP5    2.65  1.25
# 3 DP24   6.32  1.48

# GLM without intercept: test whether groups of delta fluxes are different from zero
delta.fit3 <- lm(flux ~ comp - 1, data = delta.L)
summary(delta.fit3)

# Coefficients:
#       Estimate Std. Error t value Pr(>|t|) 
# compDP0     5.642      1.620   3.482 0.002223 ** 
# compDP5     2.646      1.620   1.633 0.117352    
# compDP24    6.324      1.620   3.903 0.000819 ***

# Residual standard error: 4.583 on 21 degrees of freedom
# Multiple R-squared:  0.5885,	Adjusted R-squared:  0.5297 
# F-statistic: 10.01 on 3 and 21 DF,  p-value: 0.0002673

# Conclusion: except DP5, lactate flux increased on all resource sizes during evolution
```


## 4. Did the magnitude of flux evolution change with resource treatment? 

```{r}

# GLM with intercept: test if groups of delta fluxes are equal to each other

# Fructose:

delta.fit4 <- lm(flux ~ comp, data = delta.F)

# Perform contrasts between each resource group
contrast_DP0_DP5 <- contrast(delta.fit4, list(comp = "DP0"), list(comp = "DP5"))
contrast_DP0_DP24 <- contrast(delta.fit4, list(comp = "DP0"), list(comp = "DP24"))
contrast_DP5_DP24 <- contrast(delta.fit4, list(comp = "DP5"), list(comp = "DP24"))

#               Contrast      S.E.    Lower     Upper     t     df  Pr(>|t|)
#  DP0_DP5     5.13935    1.801056   1.393849   8.884851  2.85  21   0.0095
#  DP0_D24     1.773765   1.801056  -1.971736   5.519266  0.98  21   0.3359
#  DP5_DP24    -3.365585  1.801056  -7.111086  0.3799163 -1.87  21   0.0757

# Conclusion: evolution of fructose flux depended on resource size
# DP0 > DP5

# Acetate:

delta.fit5 <- lm(flux ~ comp, data = delta.A)

# Perform contrasts between each resource group
contrast_DP0_DP5 <- contrast(delta.fit5, list(comp = "DP0"), list(comp = "DP5"))
contrast_DP0_DP24 <- contrast(delta.fit5, list(comp = "DP0"), list(comp = "DP24"))
contrast_DP5_DP24 <- contrast(delta.fit5, list(comp = "DP5"), list(comp = "DP24"))

#               Contrast      S.E.    Lower     Upper     t     df  Pr(>|t|)
#  DP0_DP5     5.297966     2.547192  0.000790 10.59514  2.08   21     0.05
#  DP0_D24    -0.006701859  2.547192 -5.303877  5.29047  0      21   0.9979
#  DP5_DP24   -5.304668     2.547192 -10.60184 -0.00749 -2.08   21   0.0497

# Conclusion: evolution of acetate flux depended on resource size
# DP0 > DP5
# DP24 > DP5


# Lactate:

delta.fit6 <- lm(flux ~ comp, data = delta.L)

# Perform contrasts between each resource group
contrast_DP0_DP5 <- contrast(delta.fit6, list(comp = "DP0"), list(comp = "DP5"))
contrast_DP0_DP24 <- contrast(delta.fit6, list(comp = "DP0"), list(comp = "DP24"))
contrast_DP5_DP24 <- contrast(delta.fit6, list(comp = "DP5"), list(comp = "DP24"))

#               Contrast      S.E.    Lower     Upper     t     df  Pr(>|t|)
#  DP0_DP5      2.996221    2.291482 -1.769177 7.761619  1.31   21   0.2052
#  DP0_D24     -0.6818265   2.291482 -5.447224 4.083571 -0.3    21    0.769
#  DP5_DP24   -3.678047     2.291482 -8.443445 1.087351 -1.61   21   0.1234

# Conclusion: evolution of acetate flux did not depend on resource size
```

# Create metabolite flux figure (Fig. 4)

```{r}
# Start figure
png(filename="~/GitHub/Bifidobacterium/figures/Fig.4.png",
    width = 800, height = 1800, res = 96*2) 

# Set up the layout for a 3-panel figure in a single column
par(mfrow = c(3, 1),  # 3 rows, 1 column
    oma = c(6, 4, 1, 1),  # Outer margins (bottom, left, top, right)
    mar = c(2, 4, 2, 1))  # Inner margins for each plot

# Plot fructose flux
tomato1_rgb <- col2rgb("tomato")/255 

evo.dp0.F <- plot(jitter(rep(1, length(evo.DP0.F)), amount = 0.1), 
                evo.DP0.F, ylim = c(-0.5, 23), xlim = c(0.5, 3.5), 
                pch = 19, col = rgb(tomato1_rgb[1], tomato1_rgb[2], 
                tomato1_rgb[3], 0.5), cex = 1.7, yaxt = "n", xaxt = "n", 
                cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP5)
springgreen4_rgb <- col2rgb("springgreen4")/255 

points(jitter(rep(2, length(evo.DP5.F)), amount = 0.1), 
      evo.DP5.F, ylim = c(-0.5, 23), xlim = c(0.5, 3.5), pch = 19, 
      col = rgb(springgreen4_rgb[1], springgreen4_rgb[2], 
      springgreen4_rgb[3], 0.5), cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP24)
cornflowerblue_rgb <- col2rgb("cornflowerblue")/255 

points(jitter(rep(3, length(evo.DP24.F)), amount = 0.1), evo.DP24.F, 
       ylim = c(-0.5, 23), xlim = c(0.5, 3.5), pch = 19, 
       col = rgb(cornflowerblue_rgb[1], cornflowerblue_rgb[2], 
       cornflowerblue_rgb[3], 0.5),cex = 1.7, yaxt = "n", xaxt = "n", 
       cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      
box(lwd = 2)

# Add means
points(1, evo.summary.F[1,2], pch = 21, col = "white", 
       bg = "tomato1", lwd = 1, cex = 2.5)  
points(2, evo.summary.F[2,2], pch = 21, col = "white", 
       bg = "springgreen4", lwd = 1, cex = 2.5)
points(3, evo.summary.F[3,2], pch = 21, col = "white", 
       bg = "cornflowerblue", lwd = 1, cex = 2.5)

# Add SEMs
arrows(x0 = 1, y0 = evo.summary.F[1,2], 
       y1 = evo.summary.F[1,2] + evo.summary.F[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1, y0 = evo.summary.F[1,2], 
       y1 = evo.summary.F[1,2] - evo.summary.F[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 2, y0 = evo.summary.F[2,2], 
       y1 = evo.summary.F[2,2] + evo.summary.F[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 2, y0 = evo.summary.F[2,2], 
       y1 = evo.summary.F[2,2] - evo.summary.F[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 3, y0 = evo.summary.F[3,2], 
       y1 = evo.summary.F[3,2] + evo.summary.F[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

arrows(x0 = 3, y0 = evo.summary.F[3,2], 
       y1 = evo.summary.F[3,2] - evo.summary.F[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

# Add ancestral means
segments(0.5, anc.summary.F[1,1], 3.5, anc.summary.F[1,1], col = "grey", 
         lwd = 3, lty = 3, xpd = NA)

# Add axes and labels
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     labels = c("0.0", "10", "20"), at = c(0, 10, 20))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0, 10, 20), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3), labels = F)

text(2.75, 22, labels = "Fructose import", cex = 1.75)
     
# Plot acetate flux
tomato1_rgb <- col2rgb("tomato")/255 

evo.dp0.A <- plot(jitter(rep(1, length(evo.DP0.A)), amount = 0.1), 
                evo.DP0.A, ylim = c(-0.5, 32), xlim = c(0.5, 3.5), 
                pch = 19, col = rgb(tomato1_rgb[1], tomato1_rgb[2], 
                tomato1_rgb[3], 0.5), cex = 1.7, yaxt = "n", xaxt = "n", 
                cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP5)
springgreen4_rgb <- col2rgb("springgreen4")/255 

points(jitter(rep(2, length(evo.DP5.A)), amount = 0.1), 
      evo.DP5.A, ylim = c(-0.5, 30), xlim = c(0.5, 3.5), pch = 19, 
      col = rgb(springgreen4_rgb[1], springgreen4_rgb[2], 
      springgreen4_rgb[3], 0.5), cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP24)
cornflowerblue_rgb <- col2rgb("cornflowerblue")/255 

points(jitter(rep(3, length(evo.DP24.A)), amount = 0.1), evo.DP24.A, 
       ylim = c(-0.5, 32), xlim = c(0.5, 3.5), pch = 19, 
       col = rgb(cornflowerblue_rgb[1], cornflowerblue_rgb[2], 
       cornflowerblue_rgb[3], 0.5),cex = 1.7, yaxt = "n", xaxt = "n", 
       cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      
box(lwd = 2)

# Add means
points(1, evo.summary.A[1,2], pch = 21, col = "white", 
       bg = "tomato1", lwd = 1, cex = 2.5)  
points(2, evo.summary.A[2,2], pch = 21, col = "white", 
       bg = "springgreen4", lwd = 1, cex = 2.5)
points(3, evo.summary.A[3,2], pch = 21, col = "white", 
       bg = "cornflowerblue", lwd = 1, cex = 2.5)

# Add SEMs
arrows(x0 = 1, y0 = evo.summary.A[1,2], 
       y1 = evo.summary.A[1,2] + evo.summary.A[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1, y0 = evo.summary.A[1,2], 
       y1 = evo.summary.A[1,2] - evo.summary.A[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 2, y0 = evo.summary.A[2,2], 
       y1 = evo.summary.A[2,2] + evo.summary.A[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 2, y0 = evo.summary.A[2,2], 
       y1 = evo.summary.A[2,2] - evo.summary.A[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 3, y0 = evo.summary.A[3,2], 
       y1 = evo.summary.A[3,2] + evo.summary.A[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

arrows(x0 = 3, y0 = evo.summary.A[3,2], 
       y1 = evo.summary.A[3,2] - evo.summary.A[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

# Add ancestral means
segments(0.5, anc.summary.A[1,1], 3.5, anc.summary.A[1,1], col = "grey", 
         lwd = 3, lty = 3, xpd = NA)

# Add axes and labels
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     labels = c("0.0", "15", "30"), at = c(0, 15, 30))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0, 15, 30), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3), labels = F)
text(2.8, 30, labels = "Acetate export", cex = 1.75)


# Plot lactate flux
tomato1_rgb <- col2rgb("tomato")/255 

evo.dp0.L <- plot(jitter(rep(1, length(evo.DP0.L)), amount = 0.1), 
                evo.DP0.L, ylim = c(-3, 20), xlim = c(0.5, 3.5), 
                pch = 19, col = rgb(tomato1_rgb[1], tomato1_rgb[2], 
                tomato1_rgb[3], 0.5), cex = 1.7, yaxt = "n", xaxt = "n", 
                cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP5)
springgreen4_rgb <- col2rgb("springgreen4")/255 

points(jitter(rep(2, length(evo.DP5.L)), amount = 0.1), 
      evo.DP5.L, ylim = c(-3, 20), xlim = c(0.5, 3.5), pch = 19, 
      col = rgb(springgreen4_rgb[1], springgreen4_rgb[2], 
      springgreen4_rgb[3], 0.5), cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

# Plot evolved flux (DP24)
cornflowerblue_rgb <- col2rgb("cornflowerblue")/255 

points(jitter(rep(3, length(evo.DP24.L)), amount = 0.1), evo.DP24.L, 
       ylim = c(-3, 20), xlim = c(0.5, 3.5), pch = 19, 
       col = rgb(cornflowerblue_rgb[1], cornflowerblue_rgb[2], 
       cornflowerblue_rgb[3], 0.5),cex = 1.7, yaxt = "n", xaxt = "n", 
       cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      
box(lwd = 2)

# Add means
points(1, evo.summary.L[1,2], pch = 21, col = "white", 
       bg = "tomato1", lwd = 1, cex = 2.5)  
points(2, evo.summary.L[2,2], pch = 21, col = "white", 
       bg = "springgreen4", lwd = 1, cex = 2.5)
points(3, evo.summary.L[3,2], pch = 21, col = "white", 
       bg = "cornflowerblue", lwd = 1, cex = 2.5)

# Add SEMs
arrows(x0 = 1, y0 = evo.summary.L[1,2], 
       y1 = evo.summary.L[1,2] + evo.summary.L[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1, y0 = evo.summary.L[1,2], 
       y1 = evo.summary.L[1,2] - evo.summary.L[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 2, y0 = evo.summary.L[2,2], 
       y1 = evo.summary.L[2,2] + evo.summary.L[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 2, y0 = evo.summary.L[2,2], 
       y1 = evo.summary.L[2,2] - evo.summary.L[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "springgreen4")

arrows(x0 = 3, y0 = evo.summary.L[3,2], 
       y1 = evo.summary.L[3,2] + evo.summary.L[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

arrows(x0 = 3, y0 = evo.summary.L[3,2], 
       y1 = evo.summary.L[3,2] - evo.summary.L[3,3], 
       angle = 90, length = 0.1, lwd = 2, col = "cornflowerblue")

# Add ancestral means
segments(0.5, anc.summary.L[1,1], 3.5, anc.summary.L[1,1], col = "grey", 
         lwd = 3, lty = 3, xpd = NA)

# Add axes and labels
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     labels = c("0.0", "10", "20"), at = c(0, 10, 20))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0, 10, 20), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1, 2, 3), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     labels = c("DP0", "DP5", "DP24"), 
     col.axis = "black", at = c(1, 2, 3))
text(2.8, 19, labels = "Lactate export", cex = 1.75)

# Add axes titles
mtext(expression("Metabolite flux" ~ (fmol ~ cell^{-1} ~ h^{-1})), 
      side = 2, outer = TRUE, cex = 1.5, line = 1, adj = 0.5)

mtext(expression('Resource size'), side = 1, outer = TRUE, cex = 1.5, 
      line = 3.5, adj = 0.6)


# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("~/GitHub/Bifidobacterium/figures/Fig.4.png")
grid.raster(img)
```