---
title: "Probiotic evolution"
author: "Roy Moger-Reischer and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin = 2.54cm
---

# Metabolites of ancestral and evolved strains on different resources (Fig. 4)

# Clear workspace and set directory

```{r setup}
rm(list = ls())
knitr::opts_knit$set(root.dir=normalizePath("~/GitHub/Bifidobacterium"))
getwd()
```

## Load packages and functions

```{r}
require("png")
require("dplyr")
require("grid")
require("knitr")
require("extrafont")
require("ggrepel");
require("gridExtra")
require("contrast")
sem <- function(x) sqrt(var(x)/length(x))
cv <- function(x) (sd(x)/mean(x))*100
```

# Load data

```{r}
hplc <- read.csv("~/GitHub/Bifidobacterium/data/metabolite.flux.data.csv")
hplc$condition_short <- factor(hplc$condition_short,levels=c("GF","F"))
hplc$evo_trt <- factor(hplc$evo_trt, levels=c("F", "P", "X", "Anc"))
hplc$evo_trt_plotting <- recode(hplc$evo_trt, "F" = "DP0", "P" = "DP5", "X" = "DP24", "Anc" = "Ancestor")
hplc$evo_trt_plotting <- factor(hplc$evo_trt_plotting, levels = c("DP0", "DP5", "DP24", "Ancestor"))
grrPglcU <- read.csv("~/GitHub/Bifidobacterium/data/umax.on.P_F-evolved_glcU.vs.lacY.csv")
```

# Calculate metabolite fluxes 

```{r}
# Ancestral fructose flux
anc <- hplc %>%
  filter(evo_trt == "Anc")

anc.summary.F <- anc %>%
  summarize(mean = mean(flux.fru.pos, na.rm = TRUE), sem = sem(flux.fru.pos))
anc.summary.F <- as.numeric(unlist(anc.summary.F))

evo <- hplc %>%
  filter(evo_trt !="Anc")

# Evolved fructose flux
evo <- hplc %>% 
  filter(condition_short == "F" & evo_trt_plotting == "DP0") %>%
  select(Importer, flux.fru.pos, evo_trt_plotting)

evo.glcU <- evo %>%
  filter(Importer == "glcU") %>%
  select(flux.fru.pos)
evo.glcU <- as.numeric(unlist(evo.glcU))

evo.lacY <- evo %>%
  filter(Importer == "lacY") %>%
  select(flux.fru.pos)
evo.lacY <- as.numeric(unlist(evo.lacY))

evo.summary <- evo %>%
  group_by(Importer) %>%
  summarize(mean = mean(flux.fru.pos, na.rm = TRUE), sem = sem(flux.fru.pos))
evo.summary <- as.data.frame(evo.summary)
```


# Statistics: umax of glcU vs. lacY mutants (evolved on DPO) assayed on fructose (DP0)

## 1. umax of glcU vs. lacY mutants (evolved on DPO) assayed on DP5

```{r}
# test for equal variance
var.test(x = c(filter(grrPglcU, importer == "glcU")$umax_on_P), 
         y = c(filter(grrPglcU, importer == "lacY")$umax_on_P), 
         ratio = 1, alternative = "t")
# Answer: P = 0.05035; ~ homoscedastic

# Two sample t-test for differences in umax 
test.umax <- t.test(x = c(filter(grrPglcU, importer == "glcU")$umax_on_P), 
          y = c(filter(grrPglcU, importer == "lacY")$umax_on_P), mu = 0, 
          alternative = "t", paired = F, var.equal = T)
# umax of lacY mutant is 4x greater than umax of glcU mutant
# t = -4.46, df = 6, p-value = 0.004283
# mean of glcU =  0.04585;  mean of lacY = 0.18530 
```

## 2. Does fructose important differ between mutants?

```{r}
test.import <- t.test(x = evo.glcU, y = evo.lacY, mu = 0, 
          alternative = "t", paired = F, var.equal = T)

# Answer: Yes
# t = -4.0717, df = 6, p-value = 0.006563
# mean lacY = 14.514321; mean of glcU = 7.429857 
# two-fold difference in import
```

## 3. Do fructose rates differ from ancestor?

```{r}
# Subtract flux of each evolved replicate from flux of average ancestor

delta.lacY <- evo.lacY - anc.summary.F[1,1]
delta.glcU <- evo.glcU - anc.summary.F[1,1]
  
# Create data frame for analysis
delta.flux <- data.frame(flux = rbind(delta.lacY, delta.glcU), 
              comp = c(rep("lacY", nrow(delta.lacY)), rep("glcU", nrow(delta.glcU))))

# Summarize deltas
delta.summary <- delta.flux %>%
  group_by(comp) %>%
  summarize(mean = mean(flux.fru.pos, na.rm = TRUE), sem = sem(flux.fru.pos))
#   comp   mean   sem
# 1 glcU   4.72 0.640
# 2 lacY  11.8  1.62 

# GLM without intercept: test whether groups of delta fluxes are different from zero
delta.fit1 <- lm(flux.fru.pos ~ comp - 1, data = delta.flux)
summary(delta.fit1)

# Coefficients:
#         Estimate Std. Error t value Pr(>|t|)    
# compglcU    4.716      1.230   3.833  0.00863 ** 
# complacY   11.800      1.230   9.591 7.34e-05 ***

# Residual standard error: 2.461 on 6 degrees of freedom
# Multiple R-squared:  0.9468,	Adjusted R-squared:  0.929 
# F-statistic: 53.34 on 2 and 6 DF,  p-value: 0.000151

# Conclusion: fructose flux increased fro both mutants
```

## 4. Did the magnitude of flux evolution change depending on mutation? 

```{r}
delta.fit2 <- lm(flux.fru.pos ~ comp, data = delta.flux)

# Perform contrasts between each resource group
contrast_lacY_glcU <- contrast(delta.fit2, list(comp = "lacY"), list(comp = "glcU"))

#   Contrast     S.E.    Lower    Upper    t df Pr(>|t|)
# 1 7.084463 1.739931 2.827006 11.34192 4.07  6   0.0066

# Conclusion: evolution of fructose import great for lacY > glcU
```


# Figure: fructose import of glcU vs. lacY mutants

```{r}
# Start figure
png(filename="~/GitHub/Bifidobacterium/figures/Fig.5.png",
    width = 1200, height = 1200, res = 96*2) 

par(mar = c(7, 7, 5, 7))

# Plot evolved biomass (DP0)
tomato1_rgb <- col2rgb("tomato")/255 

evo.glcU.plot <- plot(jitter(rep(1.25, length(evo.glcU)), amount = 0.1), 
                evo.glcU, ylim = c(-2, 20), xlim = c(1, 2), 
                pch = 19, col = rgb(tomato1_rgb[1], tomato1_rgb[2], 
                tomato1_rgb[3], 0.5), cex = 1.7, yaxt = "n", xaxt = "n", 
                cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

points(jitter(rep(1.75, length(evo.lacY)), amount = 0.1), 
      evo.lacY, ylim = c(-2, 20), xlim = c(1, 2), pch = 19, 
      col = rgb(tomato1_rgb[1], tomato1_rgb[2], 
      tomato1_rgb[3], 0.5), cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")

box(lwd = 2)

# Add means
points(1.25, evo.summary[1,2], pch = 21, col = "white", 
       bg = "tomato1", lwd = 1, cex = 2.5)  
points(1.75, evo.summary[2,2], pch = 21, col = "white", 
       bg = "tomato1", lwd = 1, cex = 2.5) 

# Add SEMs
arrows(x0 = 1.25, y0 = evo.summary[1,2], 
       y1 = evo.summary[1,2] + evo.summary[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1.25, y0 = evo.summary[1,2], 
       y1 = evo.summary[1,2] - evo.summary[1,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1.75, y0 = evo.summary[2,2], 
       y1 = evo.summary[2,2] + evo.summary[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

arrows(x0 = 1.75, y0 = evo.summary[2,2], 
       y1 = evo.summary[2,2] - evo.summary[2,3], 
       angle = 90, length = 0.1, lwd = 2, col = "tomato1")

# Add ancestral means
segments(1.1, anc.summary.F[1], 1.9, anc.summary.F[1], col = "tomato1", 
         lwd = 3, lty = 3, xpd = NA)

# Add axes and labels
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0", "5", "10", "15", "20"), at = c(0, 5, 10, 15, 20))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(0, 5, 10, 15, 20), labels = F)
axis(side = 3, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     at = c(1.25, 1.75), labels = F)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c(expression(italic(glcU)), expression(italic(lacY))), at = c(1.25, 1.75))

mtext(expression('Evolved mutants'), side = 1,
      outer = TRUE, cex = 1.5, line = -2.5, adj = 0.5)

mtext(expression("Fructose import" ~ (fmol ~ cell^{-1} ~ h^{-1})), 
      side = 2, outer = TRUE, cex = 1.5, line = -3, adj = 0.58)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("~/GitHub/Bifidobacterium/figures/Fig.5.png")
grid.raster(img)
```